Ms. 2: Consumption and resistance across states and HRRs

### 2017-04-11 inequality-resistance/inequality-resistance.Rmd
many changes!

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, 
                      echo=TRUE, warning=FALSE, message=FALSE, cache=TRUE)
```

>>>>

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, dev=c('pdf', 'png'), 
                      echo=TRUE, warning=FALSE, message=FALSE, cache=TRUE)
import::from(broom, tidy)
```

### 2017-04-11 ms1/ms1.Rmd
many changes!

```{r setup, cache=FALSE}
# Group values in x into ranges determined by upper_limits.
# Values below the lowest limit are named as the lowest limit, and so forth.
# You should probably put Inf as the last limit.
bounds_group = function(x, upper_limits) {
  levels = lapply(seq(from=1, to=length(upper_limits)), function(i) {
    if (i == 1) {
      sprintf('< %.1f', upper_limits[i])
    } else {
      sprintf('>= %.1f and < %.1f', upper_limits[i - 1], upper_limits[i])
    }
  })
  formula_list = mapply(function(str, lim) as.formula(sprintf('x < %.1f ~ "%s"', lim, str)),
                        levels, upper_limits)
  do.call(case_when, formula_list) %>%
    factor(levels=levels, ordered=TRUE)
}

@@

multivariate_formula = paste(c('n_claims ~ age_over_65 + is_female + ', conditions), collapse=' + ') %>% formula

all_condition_model = function(usage) {
  usage %>%
    mutate(age_over_65 = age - 65) %>%
    glm(multivariate_formula, data=., family='poisson') %>%
    tidy()
}
```

@@

```{r analyze_function}
analyze = function(this_year) {
  # bene file is already filtered for plan coverage months and good geographies
  bene = read_tsv(sprintf('../bene_%i.tsv', this_year)) %>%
    filter(age >= 65, age < 116, hmo_months==0) %>%
    mutate(is_female=sex=='female', is_white=race=='white') %>%
    left_join(regions, by='state')

@@

  # then by age group
  # (the input to summarize_by is modified, since we want to group by age *group*,
  # not just by raw age)
  age_group_f = function(x) bounds_group(x, c(65, 70, 75, 80, 85, 90, 95, 116))
  claims_by_age = summarize_by(
    bene %>% mutate(age_group=age_group_f(age)),
    pde, 'age_group') %>%
    mutate(year=this_year)

@@

  claims_group_f = function(x) bounds_group(x, c(1, 2, 5, 10, 20, Inf))
  chars_by_consumption = usage %>%
    mutate(claims_group=claims_group_f(n_claims)) %>%
    group_by(claims_group) %>%
    summarize(`no. beneficiaries`=n(),
              `mean age`=mean(age),
              `% female`=mean(is_female) * 100,
              `mean no. chronic conditions`=mean(n_cc)) %>%
    mutate(year=this_year)

>>>>

```{r setup, cache=FALSE}
# Group values in x into ranges determined by upper_limits.
# Values below the lowest limit are named as the lowest limit, and so forth.
# You should probably put Inf as the last limit.
bounds_group = function(x, upper_limits) {
  levels = lapply(seq(from=1, to=length(upper_limits)), function(i) {
    if (i == 1) {
      sprintf('<= %.1f', upper_limits[i])
    } else {
      sprintf('> %.1f and <= %.1f', upper_limits[i - 1], upper_limits[i])
    }
  })
  formula_list = mapply(function(str, lim) as.formula(sprintf('x <= %.1f ~ "%s"', lim, str)),
                        levels, upper_limits)
  do.call(case_when, formula_list) %>%
    factor(levels=levels, ordered=TRUE)
}

@@

multivariate_formula = paste(c('n_claims ~ age_over_66 + is_female + ', conditions), collapse=' + ') %>% formula
start = rep(1.0, 3 + length(conditions))

all_condition_model = function(usage) {
  usage %>%
    mutate(age_over_66 = age - 66) %>%
    glm(multivariate_formula, data=., family=quasipoisson(link='identity'), start=start) %>%
    tidy()
}
```

@@

```{r analyze_function}
age_group_f = function(x) bounds_group(x, c(65, 70, 75, 80, 85, 90, 95, 100, 116))
claims_group_f = function(x) bounds_group(x, c(0, 1, 2, 5, 10, 20, Inf))

analyze = function(this_year) {
  # bene file is already filtered for plan coverage months and good geographies
  bene = read_tsv(sprintf('../bene_%i.tsv', this_year)) %>%
    filter(between(age, 66, 116), hmo_months==0) %>%
    mutate(is_female=sex=='female', is_white=race=='white') %>%
    left_join(regions, by='state')

@@

  # then by age group
  # (the input to summarize_by is modified, since we want to group by age *group*,
  # not just by raw age)
  claims_by_age = summarize_by(
      bene %>% mutate(age_group=age_group_f(age)),
      pde, 'age_group'
    ) %>%
    mutate(year=this_year)

@@

  chars_by_consumption = usage %>%
    mutate(claims_group=claims_group_f(n_claims)) %>%
    group_by(claims_group) %>%
    summarize(`no. beneficiaries`=n(),
              `mean age`=mean(age),
              `% female`=mean(is_female) * 100,
              `mean no. chronic conditions`=mean(n_cc)) %>%
    mutate(year=this_year)

### 2017-04-11 ms1/supp1.Rmd
many changes!

bene = read_tsv(sprintf('../bene_%i.tsv', this_year)) %>%
  filter(age >= 65, age < 116, hmo_months==0) %>%
  mutate(is_female=sex=='female', is_white=race=='white') %>%
  left_join(regions, by='state')

@@

usage = pde %>% count(bene_id) %>% rename(n_claims=n) %>%
  right_join(bene, by='bene_id') %>%
  replace_na(list(n_claims=0)) %>%
  left_join(cc, by='bene_id') %>%
  mutate(age_over_65 = age - 65)

@@

# All-condition negative binomial model

Each point represents a chronic condition. The coordinates are corresponding coefficients in the two models. The formula for the model is claims against age and sex.

```{r all_condition_negbin}
all_condition_formula_start = 'n_claims ~ age_over_65 + is_female + '
all_condition_formula = formula(paste(c(all_condition_formula_start, conditions), collapse=' + '))

all_condition_negbin_model = usage %>%
  MASS::glm.nb(all_condition_formula, data=.) %>%
  tidy()

all_condition_poisson_model = usage %>%
  glm(all_condition_formula, data=., family='poisson') %>%
  tidy()
```

```{r cache=FALSE}
bind_rows(mutate(all_condition_poisson_model, method='Poisson'),
          mutate(all_condition_negbin_model, method='negbin')) %>%
  mutate(condition=str_replace(term, 'TRUE$', '')) %>%
  left_join(condition_names, by='condition') %>%
  mutate(term=if_else(is.na(condition_name), condition, condition_name)) %>%
  select(term, method, estimate) %>%
  spread(method, estimate) %>%
  ggplot(aes(x=Poisson, y=negbin)) +
  geom_abline(col='gray') +
  geom_point() +
  xlim(-0.2, 0.6) + ylim(-0.2, 0.6) +
  xlab('Poisson multivariate model coefficient') +
  ylab('negative binomial multivariate model coefficient') +
  coord_fixed() +
  theme_minimal()
```

@@

```{r single_condition}
summarize_condition = function(x, condition) {
rename_(x, .dots=setNames(condition, 'has_condition')) %>%
  glm(n_claims ~ age_over_65 + is_female + has_condition, family='poisson', data=.) %>%
  tidy() %>%
  mutate(condition=condition)
}

@@

lasso_x = usage %>%
  select(age_over_65, is_female, AMI:HYPOTH) %>%
  as.matrix
lasso_y = usage$n_claims

@@

race_geography_model_f = function(condition) {
  form = formula(paste0('n_claims ~ age_over_65 + is_female + is_white + region + ', condition))
  usage %>%
    mutate(region=factor(region, levels=c('Northeast', 'West', 'Midwest', 'South'))) %>%
    glm(form, family='poisson', data=.) %>%
    tidy() %>%
    mutate(condition=condition)
}

>>>>

bene = read_tsv(sprintf('../bene_%i.tsv', this_year)) %>%
  filter(between(age, 66, 100), hmo_months==0) %>%
  mutate(is_female=sex=='female', is_white=race=='white') %>%
  left_join(regions, by='state')

@@

usage = pde %>% count(bene_id) %>% rename(n_claims=n) %>%
  right_join(bene, by='bene_id') %>%
  replace_na(list(n_claims=0)) %>%
  left_join(cc, by='bene_id') %>%
  mutate(age_over_66 = age - 66)

@@

@@

```{r single_condition}
summarize_condition = function(x, condition) {
rename_(x, .dots=setNames(condition, 'has_condition')) %>%
  glm(n_claims ~ age_over_66 + is_female + has_condition, family='poisson', data=.) %>%
  tidy() %>%
  mutate(condition=condition)
}

@@

lasso_x = usage %>%
  select(age_over_66, is_female, AMI:HYPOTH) %>%
  as.matrix
lasso_y = usage$n_claims

@@

race_geography_model_f = function(condition) {
  form = formula(paste0('n_claims ~ age_over_66 + is_female + is_white + region + ', condition))
  usage %>%
    mutate(region=factor(region, levels=c('Northeast', 'West', 'Midwest', 'South'))) %>%
    glm(form, family='poisson', data=.) %>%
    tidy() %>%
    mutate(condition=condition)
}

### 2017-04-11 states/resistance/ms2.Rmd
many changes!

@@

@@

result = bind_rows(
  state_res %>% mutate(place_type='state'),
  hrr_res %>% mutate(place_type='hrr')
)
```

@@

>>>>

spearman_model = function(df, independent_var, explanatory_vars) {
  cor.test(df[[independent_var]], df[[explanatory_vars]], method='spearman') %>% tidy
}

@@

cor_models_result = function(df) {
  group_by(df, bug, drug_group) %>%
    do(spearman_model(., 'mean_percent_nonsusceptible', 'cpkp')) %>%
    ungroup %>%
    #filter(term=='cpkp') %>%
    mutate(adj.p.value=p.adjust(p.value, method='BH'),
           sig=pvalue_stars(adj.p.value))
}

@@

result = bind_rows(
  state_res %>% mutate(place_type='state'),
  hrr_res %>% mutate(place_type='hrr')
)

cor_result = bind_rows(
  cor_models_result(state_dat) %>% mutate(place_type='state'),
  cor_models_result(hrr_dat) %>% mutate(place_type='hrr')
)
```

@@

## Cor vs lm

I asked whether using Spearman correlations is better/different than using linear models. For the most part, they identify the same relationships.

```{r cor_vs_lm}
f = function(x) select(x, bug, drug_group, adj.p.value, place_type)
x = bind_rows(
  cor_result %>% f %>% mutate(model='cor'),
  result %>% f %>% mutate(model='lm')
) %>% spread(model, adj.p.value)

x %>% filter(place_type=='state') %>% ggplot(aes(x=lm, y=cor)) +
  geom_point() +
  scale_x_log10() + scale_y_log10() +
  geom_abline() + geom_hline(yintercept = 0.05) + geom_vline(xintercept = 0.05) +
  coord_fixed()

x %>% filter(cor < 0.05, lm > 0.05) %>% kable
```
